---
title: "Table theme gallery"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

Compare the themes on two reusable tables: a compact example with four rows
and two spanners, and a complete example with row groups, subtotals, a grand
total, footnotes, and a source note. The examples adapt the cars and pizza
sales prototypes and use datasets bundled with **gt**.

Build the table first, including summaries and notes, then apply the theme.
All Hokusai examples use `stripe = TRUE` and `gridlines = FALSE`: only the
palette changes. The shared neutral stripes stay the same; look at the title,
section rules, group headings, and summary fills to compare colors.

## Build the examples

```{r libraries}
library(ekiotable)
library(gt)
```

### Compact table

Four cars, a simple title, row labels, and spanners for performance and fuel
economy keep the structure small enough to compare at a glance.

```{r compact-table}
cars_data <- gtcars[order(gtcars$msrp, decreasing = TRUE), ]
cars_data <- cars_data[
  seq_len(4),
  c("model", "hp", "trq", "mpg_c", "mpg_h")
]

simple_table <- cars_data |>
  gt(rowname_col = "model") |>
  tab_header(title = "High-performance cars") |>
  tab_stubhead(label = "Model") |>
  tab_spanner(label = "Performance", columns = c(hp, trq)) |>
  tab_spanner(label = "Fuel economy (mpg)", columns = c(mpg_c, mpg_h)) |>
  cols_label(
    hp = "HP",
    trq = "Torque (lb-ft)",
    mpg_c = "City",
    mpg_h = "Highway"
  ) |>
  fmt_integer(columns = c(hp, trq, mpg_c, mpg_h))
```

### Complete table

Aggregate pizza sales by category and size. Each dataset row represents one
pizza sold, so the count below measures pizzas rather than distinct orders.
All sizes are retained, including XL and XXL, so the grand total covers the
entire dataset.

```{r complete-data}
pizza_source <- data.frame(
  category = gt::pizzaplace$type,
  size = gt::pizzaplace$size,
  pizzas = 1L,
  revenue = gt::pizzaplace$price
)
pizza_data <- aggregate(
  cbind(pizzas, revenue) ~ category + size,
  data = pizza_source,
  FUN = sum
)
pizza_data$average_price <- pizza_data$revenue / pizza_data$pizzas
pizza_data$size <- factor(
  pizza_data$size,
  levels = c("S", "M", "L", "XL", "XXL")
)
pizza_data <- pizza_data[order(pizza_data$category, pizza_data$size), ]
```

The category column defines row groups (`groupname_col`), and size supplies
the stub labels (`rowname_col`). Only additive quantities receive totals;
average prices are left blank in summary rows.

```{r complete-table}
complete_table <- pizza_data |>
  gt(rowname_col = "size", groupname_col = "category") |>
  tab_header(
    title = "Pizza sales by category and size",
    subtitle = "Full-year sales, 2015"
  ) |>
  tab_stubhead(label = "Category / size") |>
  tab_spanner(label = "Sales", columns = c(pizzas, revenue)) |>
  cols_label(
    pizzas = "Pizzas sold",
    revenue = "Revenue",
    average_price = "Average price"
  ) |>
  fmt_integer(columns = pizzas) |>
  fmt_currency(columns = c(revenue, average_price), currency = "USD") |>
  summary_rows(
    groups = everything(),
    columns = pizzas,
    fns = list(Total = ~ sum(.)),
    fmt = ~ fmt_integer(.)
  ) |>
  summary_rows(
    groups = everything(),
    columns = revenue,
    fns = list(Total = ~ sum(.)),
    fmt = ~ fmt_currency(., currency = "USD")
  ) |>
  grand_summary_rows(
    columns = pizzas,
    fns = list("Grand total" = ~ sum(.)),
    fmt = ~ fmt_integer(.)
  ) |>
  grand_summary_rows(
    columns = revenue,
    fns = list("Grand total" = ~ sum(.)),
    fmt = ~ fmt_currency(., currency = "USD")
  ) |>
  tab_footnote(
    footnote = "Counts pizzas sold, not distinct orders; an order can contain several pizzas.",
    locations = cells_column_labels(columns = pizzas)
  ) |>
  tab_footnote(
    footnote = "Revenue divided by pizzas sold within each category and size.",
    locations = cells_column_labels(columns = average_price)
  ) |>
  tab_source_note(
    source_note = md("Source: **gt::pizzaplace**, pizza sales in 2015.")
  )
```

## EKIO

`gt_theme_ekio()` uses filled blue column headers and spanners, unfilled group
headings with Baltic Blue rules, and a dark grand total. It does not add a
footer.

::: {data-bs-theme="light"}

```{r ekio-simple}
simple_table |> gt_theme_ekio(stripe = TRUE)
```

:::

::: {data-bs-theme="light"}

```{r ekio-complete}
complete_table |> gt_theme_ekio(stripe = TRUE)
```

:::

## Hokusai palettes

Hokusai uses white headers, fine rules, and pale summary backgrounds.
Typography, striping, and gridlines are held constant across the four palettes.
These are theme colors, not a scale encoding the values in the table.

### Mountain

Deep blue titles and accents with a nearly white, warm summary tint.

::: {data-bs-theme="light"}

```{r mountain-simple}
simple_table |>
  gt_theme_hokusai(palette = "mountain", stripe = TRUE, gridlines = FALSE)
```

:::

::: {data-bs-theme="light"}

```{r mountain-complete}
complete_table |>
  gt_theme_hokusai(palette = "mountain", stripe = TRUE, gridlines = FALSE)
```

:::

### Wind

Muted slate-blue titles and accents with a warm gray summary tint.

::: {data-bs-theme="light"}

```{r wind-simple}
simple_table |>
  gt_theme_hokusai(palette = "wind", stripe = TRUE, gridlines = FALSE)
```

:::

::: {data-bs-theme="light"}

```{r wind-complete}
complete_table |>
  gt_theme_hokusai(palette = "wind", stripe = TRUE, gridlines = FALSE)
```

:::

### Blossom

Dark blue titles, a separate blue accent, and a pale beige summary tint.

::: {data-bs-theme="light"}

```{r blossom-simple}
simple_table |>
  gt_theme_hokusai(palette = "blossom", stripe = TRUE, gridlines = FALSE)
```

:::

::: {data-bs-theme="light"}

```{r blossom-complete}
complete_table |>
  gt_theme_hokusai(palette = "blossom", stripe = TRUE, gridlines = FALSE)
```

:::

### Lake

Slate-blue titles, a lighter blue accent, and a pale cream summary tint.

::: {data-bs-theme="light"}

```{r lake-simple}
simple_table |>
  gt_theme_hokusai(palette = "lake", stripe = TRUE, gridlines = FALSE)
```

:::

::: {data-bs-theme="light"}

```{r lake-complete}
complete_table |>
  gt_theme_hokusai(palette = "lake", stripe = TRUE, gridlines = FALSE)
```

:::
